Formal verification of train route planning

### 2025-06-27 paper/main.tex

The system is then represented by a formula:
\begin{verbatim}
    Spec == Init /\  [][Next]_variables
\end{verbatim}%Revoir la syntaxe au cas où / remplacer par le jolie tla2latex
where variables is the tuple of all variables. The intended meaning is that the initial configuration is specified in \mintinline{C}{Init} and steps follow the \mintinline{C}{Next} formula.  Actually, presently, in our formal semantics, nothing forces a rule to be applied. As a consequence, an ongoing train may never reach its next block. Thanks to TLA+, we can ensure that a rule will eventually be applied (if it can be). Those are known as weak fairness conditions. Again, we specify it as follows. \todo{WF condition?}

@@

\paragraph{Experiments}  With TLA+, we checked that any safe step is actually (not surprisingly!) reliable.  On our network, one finds 58 368 safe steps, all being reliable. Checking all the paths took approximately 30 hours on an Linux desktop machine with an Intel i5-4590 (4-cores at 3.7GHz) and 8GB of memory.

@@

\subsection{Simplification}

\subsection{Experiments}

@@

Parler des expériences

>>>>

The system is then represented by a formula:
\begin{verbatim}
    Spec == Init /\ [][Next]_variables /\ WF_variables(Next)
\end{verbatim}%Revoir la syntaxe au cas où / remplacer par le jolie tla2latex
where variables is the tuple of all variables. The intended meaning is that the initial configuration is specified in \mintinline{C}{Init} and steps follow the \mintinline{C}{Next} formula.  Without the weak fairness condition \mintinline{C}{WF_variables(Next)},  nothing forces a rule to be applied. As a consequence, an ongoing train could never reach its next block. The weak fairness condition prevents this incorrect simulation.

@@

\paragraph{Experiments}  With TLA+, we checked that the 58 368 safe steps are actually (not surprisingly!) correct. Checking all the paths took approximately 30 hours on an Linux desktop machine with an Intel i5-4590 (4-cores at 3.7GHz) and 8GB of memory.

@@

\paragraph{Optimisations. } We apply some optimisations on the orders. For instance, we remove orders that do not alternate: $\su{d}{\vec{b}} \cdot \su{d}{\vec{b}} \cdot P$ by $\su{d}{\vec{b} \cdot \vec{b'}} \cdot P$. In the same way,  we remove empty lists, singleton lists: if $Len(\vec{b}) \leq 1$, we replace $\su{d}{\vec{b}} \cdot P$ by $P$. We have few more heuristics that we do not show here due to lack of space.

\paragraph{Experiments}  Parler des expériences

@@
